Accuracy and precision of zero-heat-flux temperature measurements with the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System

### 2020-05-06 manuscript/R1/response/response.Rmd

---
toname: Journal of Clinical Monitoring and Computing
toaddress: Editorial Board
opening: To the editor, 
closing: Regards,
spacing: 1
output: ConwayLetterTemplate::letter
---

Thank you for allowing us to revise our manuscript for further consideration for publication. Please find our responses to the comments from the reviewers below.

# Additional changes

- Title changed to: "Accuracy and precision of zero-heat-flux temperature measurements with the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System: A systematic review and meta-analysis"

# Reviewer 1

## General comments

*Without the previous Figure 2, which has been removed likely because it was a 100MB large image instead of a Table, there is no overview table of the included studies present in this paper and the reader has to visit the authors website to obtain this critical piece of information. I don't find this acceptable for a systematic review, as critical pieces are not contained in the publication and the preservation of the website can't be guaranteed.*

Response:

>We have created a table to describe the study characterstics to be included in the main manuscript (Table 2).

*The website seems to be out of sync with the paper, as references [6, 23-37], [26,38], and [39] are indeed 18 studies, but the figure at "Graphical overview of the findings", which shows 16 studies does not agree with the list of references - West 2019 [37] and Iden 2015 [28] are both missing from the plot. Subsequently I question the total numbers of patients and samples in the remainder of the text.*

Response:

>The 'Graphical overview of findings' figures in the website present the results of each meta-analysis that was undertaken. The primary meta-analysis contains 16 studies because Iden and West were not included in this particular meta-analysis, as they did not report results of a comparison between ZHF and core temperature. They used nasopharyngeal or sublingual. With the new Table 2 to be included in the main manuscript, we have included the study characteristics for these two studies though.

## Specific comments

*Abstract, P2, L16: I am not sure if "invasive" is the best word for a nasopharangeal or oropharangeal temperature probe compared to let's say a pulmonary artery blood temperature? Maybe "more invasive"?*

Response:

>Changed to 'more' invasive.

*Introduction, P6, L36: I believe the system is now referred to as the "3M Bair Hugger Temperature Monitoring System", although still widely referred to as SpotOn?*

Response:

>We have made this distinction more clear in the introduction. It now reads:

>"Originally developed in the 1970s, zero-heat-flux technology was recently implemented in the 3M Bair Hugger Temperature Monitoring System (3M, St Paul, MN), as a single-use, disposable sensor. This device was previously known as the SpotOn Temperature Monitoring System. Another device on the market is the T-core sensor from Dräger."

*Methods, P9, L9: While the shiny-based website is really neat, it might be advisable to have the search strategy included as an Appendix to the manuscript to ensure it will remain available in the figure?*

Response:

>Included the search strategy as an appendix for the main manuscript.

*Methods, P11, L40: Could you at least get a DOI to archive the code/files to provide a stable place to refer to when this Github repository becomes unavailable?*

Response:

>The github repo containing data and code has been added to Zenodo to attain a DOI.

*Methods P11, L53 and P12, L7: These are kind of mixing results and methods; which might be okay here, but you might want to provide the references for these studies?*

Response:

>We have now included the refernces.

*Methods, P12, L54: Did you consider to do subgroup analyses by measurement site? (Or by population age - pediatric vs. adult)*

>We did undertake.......

*Results, P13, L20: Please cite all studies used here. Also I am not convinced that a website link to contain one of the key elements for any systematic review: the overview table, is acceptable instead of converting the figure into a table and including it in the paper.*

Response:

>We have created a table to describe the study characterstics to be included in the main manuscript (Table 2). Citations were added to the text.

*Results, P13, L27: These values don't agree with the website. Figure: Participants 827 vs. text: participants 756; also you missed the West 2019 paper here, so this needs updating.*

Response:

>The numbers of participants reported in the manuscript is correct. The study by Boisson et al. (2018) split the analyses in their study into times during surgery where there were 'rapid' core temperature changes and 'slow' core temperature changes. As such, the total number of participants reported in the website (n=827) counted the number of participants in the Boisson et al. (2018) study twice. We added that explanation to the website under the figure.

*Results, P13, L28: Which 16 studies/8 studies etc?*

Response:

>All studies included in the review have been cited at least once in this introductory paragraph for the results. We believe there is little value in citing each of the studies included in each of the meta-analyses, as this information is better presented in the new Table 2 and in the graphical overview of findings figures in the website.

*Results, P13, L57: The number of subjects can't be the number of measurements, something went wrong here when updating the paper. (You report n=756 in the abstract.)*

Response:

>Thank you for noticing this mistake. The number of participants in this meta-analysis for the nasopharyngeal subgroup was corrected.

*Results, P14, L15: I point this out in my comment in the introduction, why do you clarify it in the results?*

Response:

>Thank you for the suggestion. We moved the explanation to the introduction.

*Results, P14, L35: Can you at least refer to the supplemental material here, if you are not planning to include the bias assessment in the paper?*

Response:

>We have added one of the risk of bias figures to the manuscript (Figure 2) to show risk of bias judgements for each of the included studies.

*Results, P16, L48: So why are you telling us about it if you then don't report it?*

Response:

>

*Discussion, P19, L14: One would assume that the PICU study (Idei) is also in children, not only Carvalho?*

Response:

>Thank you for noticing this mistake - text has been corrected to 'two studies'.

*Discussion, P19, L43: But isn't clinical utility important? Do you want to elaborate here?*

Response:

>

*References: #5 seems incomplete; #7 seems incomplete; #11 is very incomplete, #21 is incomplete; #27 is incomplete*

Response:

>

*Figure 1: Isn't "2 studies awaiting classification" a bit unusual? - You excluded them as they are available only in abstract form, right?*

Response:

>The studies awaiting classification approach is used by Cochrane. In future updates to the review these studies reported in abstract form may be published and full and we could then include them.

# Reviewer 2

*This is a timely and really worthwhile meta-analysis. My main concern is that the level of evidence - moderate at best - is insufficient to draw any definitive conclusions about the utility of the method. It is also worth noting that this method is an 'indirect measurement' as opposed to peripheral thermometers which give an 'indirect estimate' i.e. the temperature they display is not the temperature they read (see the latest NICE CG65 for more detail)*

Response:

>

*I would like to see a figure for the proportion of readings within the acceptable range.*

Response:

>Although some (not all) studies may have reported this outcome, we believe the statistical approach we used to combine results from individual studies is a more robust method to estimate the accuracy of this temperature monitoring device. This is because as it is able to account for potentially unadjusted repeated measures estimates as well as incorporate the degree of inconsistency between results and sampling error into the estimates of the population limits of agreement. In addition, the '% within acceptable limits' outcome is not a very useful outcome for application to practice, as one can never know which measurement was within or not within the limits. With the population limits of agreement approach, we can interpret that each measurement taken from one of these devices may be as much as 1 degree higher or lower than 'core' temperature.

*Most of all, I think the best (and only really acceptable) conclusion from this paper would be to define the exact characteristics of a trial which would determine the accuracy of ZF thermometers with a high level of evidence.  I also think that such a trial would need to be carried out for every device separately because there are small but important differences in the construction of each of them.*

Response:

>

*The introduction should be reviewed and made to read more clearly.  Repetition of the word "commonly' in consecutive sentences for example..*

Response:

>We have removed the repetition of the 'commonly' term.

*The description of the mechanism of action of the ZFT does not clearly explain it to someone who doesn't already understand it.*

Response:

>

*13/57 I'm sure there were 109k measurement but there weren't that number of participants*

Response:

>Thank you for noticing this mistake. We have corrected it to the number of participats.

### 2020-05-06 manuscript/manuscript.Rmd
started R1 for jcmc

---
title: "Accuracy and precision of zero-heat-flux temperature monitoring"
subtitle: A systematic review and meta-analysis
# description: "A systematic review and meta-analysis"
# author:
# - name: Aaron Conway
#   affiliation: Peter Munk Cardiac Centre, UHN & Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
#   email: aaron.conway@utoronto.ca
#   url: https://www.aaronconway.info
# - name: Megan Bittner
#   affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
# - name: Dan Phan
#   affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
# - name: Navpreet Kamboj
#   affiliation: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto
#   email: navpreet.kamboj@mail.utoronto.ca
# - name: Kristina Chang
#   affiliation: Toronto General Hospital, UHN
# - name: Elizabeth Tipton
#   affiliation: Department of Statistics, Northwestern University
# - name: Matteo Parotto
#   affiliation: Toronto General Hospital, UHN
# date: "`r Sys.Date()`"
# repository_url: https://github.com/awconway/zhf-review
# twitter:
#   creator: "@aw_conway"
# output: distill::distill_article
bibliography: Extras/bibliography.bib
csl: Extras/sp-basic.csl
author:
  - Aaron Conway PhD:
      email: aaron.conway@utoronto.ca
      phone: +1 416 946 7112
      address: 585 University Ave | Toronto, ON  M5G 2N2
      institute: [PMCC, LBFON, QUT]
      correspondence: true
  - Megan Bittner BSc:
      institute: [LBFON]
      correspondence: false
  - Dan Phan BSc:
      institute: [LBFON]
      correspondence: false
  - Kristina Chang MScN:
      institute: [PMCC]
      correspondence: false
  - Navpreet Kamboj BScN:
      institute: [LBFON]
      correspondence: false
  - Elizabeth Tipton PhD:
      institute: [NW]
      correspondence: false
  - Matteo Parotto MD, PhD:
      institute: [TGH, UoT]
      correspondence: false
institute:
  - PMCC: Peter Munk Cardiac Centre, University Health Network, Toronto, Canada
  - LBFON: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto, Toronto, Canada
  - QUT: School of Nursing, Queensland University of Technology (QUT), Brisbane, Australia
  - NW: Department of Statistics, Northwestern University, Evanston, IL, United States
  - TGH: Department of Anesthesia and Pain Management, Toronto General Hospital, UHN, Toronto, Canada
  - UoT: Department of Anesthesia and Interdepartmental Division of Critical Care Medicine, University of Toronto, Toronto, Canada
output:
  bookdown::word_document2:
    reference_docx: Extras/manuscript_template.docx
    pandoc_args:
      - --lua-filter=Extras/scholarly-metadata.lua
      - --lua-filter=Extras/author-info-blocks.lua
always_allow_html: true
params:
  distill: FALSE
  word: TRUE
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(include=FALSE,
                      echo=FALSE, warning=FALSE, message=FALSE)
library(dplyr)
library(ggprisma) # https://github.com/awconway/ggprisma
library(kableExtra)
library(knitr)
library(magrittr)
source("R/loa_maker.R")
source("R/meta.R")
```

@@

<div class="l-page-outset">
<blockquote>
### Abstract

@@

Zero-heat-flux thermometers provide clinicians with the ability to continuously and non-invasively monitor body temperature. These devices are increasingly being used to substitute for invasive core temperature measurement during surgery and in critical care. The aim of this review was to determine the accuracy and precision of zero-heat-flux temperature measurements.

@@

The range of uncertainty in the accuracy of the zero-heat-flux thermometer should be taken into account when using this device to inform clinical decision-making. Clinicians should therefore consider the potential that a temperature measurement from a zero-heat-flux thermometer could be as much as 1°C higher or lower than core temperature. Use of this device may not be appropriate in situations where a difference in temperature of less than 1°C is important to detect.

@@

</blockquote>
</div>

@@

```{r conditional_title_page, results='asis', include=params$word}
cat("
## Declarations

Availability of data and material (data transparency): All data used in the meta-analyses is available [here](https://aconway.dev/zhf).

@@

")
```

@@

Continuous core temperature monitoring is commonly used within perioperative and intensive care settings. Thermoregulatory dysfunction is commonly associated with the induction of anesthesia and can lead to adverse outcomes including cardiac arrhythmia, altered hemostasis and increased risk of surgical site infection.[@frank1997perioperative; @kurz1996perioperative; @michelson1994reversible; @rohrer1992effect] The pulmonary artery catheter is the reference standard for continuous core temperature monitoring, however, the invasive nature of this method renders an increased risk of bloodstream infections and damage to surrounding tissue.[@hadian2006evidence] Although clinically accurate relative to pulmonary artery temperature measurements, surrogate measures of core temperature at esophageal, rectal, and bladder locations remain mildly invasive interventions.

Zero-heat-flux thermometry is an alternative non-invasive method that allows for continuous monitoring of core temperature. Originally developed in the 1970s, zero-heat-flux technology was recently implemented in the 3M SpotOn temperature monitoring system (3M, St Paul, MN), as a single-use, disposable sensor. In practice, the zero-heat-flux sensor, such as the 3M SpotOn, is placed on the lateral surface of the forehead and is initially warmed to equilibrate the temperature of the skin surface to the underlying core tissues. Equipped with a thermal insulator, the zero-heat-flux sensor eliminates heat loss to the environment to allow for changes in core temperature to be directly reflected by a change in skin surface temperature.[@eshraghi2014]

The agreement between zero-heat-flux thermometers and core as well as other peripheral thermometers has been investigated in multiple studies over the last five years. Appraisal of these studies and synthesis of the results would aid clinicians in deciding the appropriate circumstances in which zero-heat-flux thermometers may be used. We aimed to determine if zero-heat-flux thermometers have clinically acceptable accuracy and precision relative to established core and peripheral temperature measurement devices. Accuracy is defined as the average difference between temperature measurements from the zero-heat-flux and comparator device and precision as the variance (standard deviation) in the differences.

@@

A systematic review was conducted in accordance with a predetermined protocol. The primary comparison for this review was temperature measured from a zero-heat-flux thermometer versus temperature measured from a core site, which we defined as temperature measured at either an arterial, esophageal, bladder or rectal site. Secondary comparisons were made between zero-heat-flux thermometers and temperatures taken at peripheral sites.

@@

Observational studies that reported temperature measurements from a zero-heat-flux thermometer and comparator thermometer were included. Studies involving a case control design were excluded due to potential for overestimation of the intervention performance. Studies were excluded if conducted on non-human subjects or outside of a clinical healthcare setting. No publication date restrictions were applied. Published conference abstracts were included if there was enough information reported to appraise the quality of the study. There were no language restrictions applied during the search.

@@

Published studies were found by searching Medline and EMBASE to January 2020. The Cochrane-recommended search strategy combining terms for the ‘target condition’ and ‘index test’ was used.[@mann2012should] This search strategy is an efficient approach for systematic reviews of diagnostic test accuracy studies.[@preston2015improving] We also conducted forward citation searching, by using Google Scholar to search the citations of the first article published on the accuracy of zero-heat-flux thermometers.[@eshraghi2014] The search strategies used for each database are available on a [webpage](https://aconway.dev/zhf) containing supplementary information for the systematic review. Selection of studies was undertaken independently by two reviewers.

@@

The objective for the meta-analysis was to estimate the population limits of agreement between temperature measurements from the zero-heat-flux and established comparator thermometers. A framework for meta-analysis of Bland-Altman method comparison studies based on limits of agreement approach was used.[@tipton2017framework]  This method was selected because it parallels the approach used in primary Bland-Altman studies, whereby an estimate is generated for the pooled limits of agreement in the population (not just in the samples studied). The ‘population limits of agreement’ is more broad than the limits of agreement commonly reported in the meta-analyses of Bland-Altman studies.[@tipton2017framework] Pooled limits of agreement were calculated using $\delta\pm2\sqrt{\sigma2+\tau2}$, where $\delta$ is the average bias across studies, $\sigma2$ is the average within-study variation in differences and $\tau2$ is the variation in bias across studies.

@@

Prior to conducting the meta-analyses, the results from each study were converted into a standard format, with bias meaning comparator thermometer minus zero-heat-flux thermometer measured in degrees Celsius (°C). In several studies, results were reported for multiple groups of participants, therefore in the meta-analysis each of these groups was treated as a separate 'comparison'. Other studies reported multiple sets of results, whereby analyses were conducted between zero-heat-flux and various comparator devices used on the same participant. These instances were also treated as a separate 'comparison' if the comparator devices were a part of separate meta-analysis groups. One study reported intraoperative, postoperative and overall results for the same participants. Only the paired measurements from the overall results were included in the main and low risk bias analyses, leaving paired measurements exclusively from the intraoperative and postoperative timepoints to be included in respective meta-analyses subgroups.

@@

Eighteen studies were included (Figure 1). Two studies reported only in abstract form were not included and assigned as 'studies awaiting classification' because there was insufficient information provided.[@cullen2015measuring; @VendrellJorda2016]

```{r fig-1,  fig.cap="PRISMA Flow Diagram", echo=FALSE, fig.width=5.5, fig.path="figures/", include=params$distill, eval=params$distill}
# https://github.com/awconway/ggprisma
ggprisma::ggprisma(retrieved = 131, included = 18, duplicates = 35, full_text = 23, wrong_intervention = 3, wrong_comparator = 2, wrong_design = 1, awaiting_classification = 2)
```

Detailed characteristics of included studies can be viewed on the [webpage](https://aconway.dev/zhf/) containing supplementary information for the systematic review. The primary comparison of zero-heat-flux versus core temperature measurements (eg, arterial, bladder, esophageal or rectal) consisted of `r primary$Comparisons` comparisons from `r primary$Studies` individual studies. In total, data from `r primary$Participants` participants with `r format(primary$Measurements, scientfic=F)` paired measurements were included in this comparison (two studies did not report the total number of measurements included in their analysis). The sensitivity analysis for the primary comparison with only studies that were judged as low risk of bias across all domains included `r low_risk$Comparisons` comparisons from `r low_risk$Studies` studies that enrolled `r low_risk$Participants` participants with `r low_risk$Measurements` paired measurements (two studies did not report the total number of measurements included in their analysis). There were `r ICU$Studies` studies that compared zero-heat-flux to core temperature measurements in ICU patients, comprising `r ICU$Comparisons` comparisons with `r ICU$Measurements` measurements from `r ICU$Participants` participants. Another `r OT$Studies` studies were included that compared zero-heat-flux to core temperature measurements in patients undergoing surgery, comprising `r OT$Comparisons` comparisons with `r OT$Measurements` measurements from `r OT$Participants` participants. Nasopharyngeal thermometers were used as the comparator devices in `r NPA$Studies` studies, with `r NPA$Measurements` paired measurements from `r NPA$Measurements` participants. Sublingual thermometers were used as the comparator devices in `r SL$Studies` studies that reported `r SL$Measurements` paired measurements from `r SL$Participants` participants.

Most studies included in this systematic review evaluated the zero-heat-flux temperature monitoring system manufactured by 3M.[@Boisson_2018; @Carvalho_2019; @Dahyot_Fizelier_2017; @eshraghi2014; @G_mez_Romero_2019; @idei2016; @Iden_2015; @Jack_2019; @Kato_2015; @kollmann2019validation; @M_kinen_2016; @Morettini_2019; @Pesonen_2018; @Schell_Chaple_2018; @Tachibana_2019; @West_2019] Previously known as the SpotOn Temperature Monitoring System, the 3M zero-heat-flux device is now referred to commercially as the Bair Hugger Temperature Monitoring System. Two studies evaluated the Tcore^TM^ device.[@sastre2019evaluation; @G_mez_Romero_2019] The oldest study included in this review, published in 2000, was an evaluation of the Coretemp device from Terumo.[@Harioka_2000] All comparisons reported adherence to the zero-heat-flux device manufacturer instructions and placed the sensor on the forehead of the participants. One included study also reported results for comparisons where the zero-heat-flux thermometer was not placed on the forehead. We did not include these comparisons in our meta-analysis.

High risk of bias was associated with patient selection for 13 (43%) comparisons from 9 (56%) studies, conduct of zero-heat-flux and comparator measurements in 10 (33%) comparisons from 8 (50%) studies and 12 (40%) comparisons 9 (30%) studies, respectively (mostly due to zero-heat-flux measurements being taken with knowledge of the comparator measurements and vice versa) and participant flow for 13 (43%) studies. In 19 (50%) comparisons from 8 (50%) studies, the authors had declared a conflict of interest or receipt of funding or equipment from the manufacturer of the zero-heat-flux device under evaluation.

### Primary comparison: zero-heat-flux thermometer versus core thermometers

Table 1 presents results of the primary, sensitivity and all subgroup analyses for the primary comparison as well as secondary comparisons. The pooled estimate for the mean bias between zero-heat-flux and core temperature measurements was `r primary$bias_mean`°C. Population limits of agreement, which take into consideration the between-study heterogeneity and sampling error, were wide, spanning from `r primary$CI_L_rve`°C to `r primary$CI_U_rve`°C (`r primary$Measurements` measurements; `r primary$Participants` participants; `r primary$Studies` studies). The amount of between-study heterogeneity is displayed graphically in the density plot in Figure 2. The quality of evidence for the primary comparison was downgraded to low quality due to concerns about study limitations and inconsistency. Population limits of agreement for the sensitivity analysis restricted to studies rated as having low risk of bias across all the domains of the QUADAS-2 were similar to the primary analysis (`r low_risk$Measurements` measurements; `r low_risk$Participants` participants; `r low_risk$Studies` studies). The mean bias was again `r low_risk$bias_mean`°C with population limits of agreement spanning from `r low_risk$CI_L_rve`°C to `r low_risk$CI_U_rve`°C. A further sensitivity analysis excluding studies that received industry funding revealed population limits of agreement that were considerably wider than the primary analysis (`r conflict$CI_L_rve`°C to `r conflict$CI_U_rve`°C). The evidence rating for this sensitivity analysis was downgraded to low quality though, due to concerns about study limitations, inconsistency and imprecision.

>>>>

---
title: "Accuracy and precision of zero-heat-flux temperature measurements with the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System"
subtitle: A systematic review and meta-analysis
bibliography: Extras/bibliography.bib
csl: Extras/sp-basic.csl
author:
  - Aaron Conway PhD:
      email: aaron.conway@utoronto.ca
      phone: +1 416 946 7112
      address: 585 University Ave | Toronto, ON  M5G 2N2
      institute: [PMCC, LBFON, QUT]
      correspondence: true
  - Megan Bittner BSc:
      institute: [LBFON]
      correspondence: false
  - Dan Phan BSc:
      institute: [LBFON]
      correspondence: false
  - Kristina Chang MScN:
      institute: [PMCC]
      correspondence: false
  - Navpreet Kamboj BScN:
      institute: [LBFON]
      correspondence: false
  - Elizabeth Tipton PhD:
      institute: [NW]
      correspondence: false
  - Matteo Parotto MD, PhD:
      institute: [TGH, UoT]
      correspondence: false
institute:
  - PMCC: Peter Munk Cardiac Centre, University Health Network, Toronto, Canada
  - LBFON: Lawrence S. Bloomberg Faculty of Nursing, University of Toronto, Toronto, Canada
  - QUT: School of Nursing, Queensland University of Technology (QUT), Brisbane, Australia
  - NW: Department of Statistics, Northwestern University, Evanston, IL, United States
  - TGH: Department of Anesthesia and Pain Management, Toronto General Hospital, UHN, Toronto, Canada
  - UoT: Department of Anesthesia and Interdepartmental Division of Critical Care Medicine, University of Toronto, Toronto, Canada
output:
  officedown::rdocx_document:
    reference_docx: Extras/manuscript_template.docx
    pandoc_args:
      - --lua-filter=Extras/scholarly-metadata.lua
      - --lua-filter=Extras/author-info-blocks.lua
always_allow_html: true
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(include=TRUE,
                      echo=FALSE, warning=FALSE, message=FALSE)
library(dplyr)
library(ggprisma) # https://github.com/awconway/ggprisma
library(kableExtra)
library(knitr)
library(magrittr)
source(here::here("manuscript", "loa_maker.R"))
source(here::here("manuscript","meta.R"))
library(gofer) # https://github.com/awconway/gofer
library(flextable)
```

@@

### Abstract

@@

Zero-heat-flux thermometers provide clinicians with the ability to continuously and non-invasively monitor body temperature. These devices are increasingly being used to substitute for more invasive core temperature measurements during surgery and in critical care. The aim of this review was to determine the accuracy and precision of zero-heat-flux temperature measurements from the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System.

@@

The range of uncertainty in the accuracy of the zero-heat-flux thermometer should be taken into account when using this device to inform clinical decision-making. Clinicians should therefore consider the potential that a temperature measurement from a 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System could be as much as 1°C higher or lower than core temperature. Use of this device may not be appropriate in situations where a difference in temperature of less than 1°C is important to detect.

@@

@@

## Declarations

Availability of data and material (data transparency): All data used in the meta-analyses is available [here](https://zhf-info.netlify.app).

@@

@@

Continuous core temperature monitoring is commonly used within perioperative and intensive care settings. Thermoregulatory dysfunction is associated with the induction of anesthesia and can lead to adverse outcomes including cardiac arrhythmia, altered hemostasis and increased risk of surgical site infection.[@frank1997perioperative; @kurz1996perioperative; @michelson1994reversible; @rohrer1992effect] The pulmonary artery catheter is the reference standard for continuous core temperature monitoring, however, the invasive nature of this method renders an increased risk of bloodstream infections and damage to surrounding tissue.[@hadian2006evidence] Although clinically accurate relative to pulmonary artery temperature measurements, surrogate measures of core temperature at esophageal, rectal, and bladder locations remain mildly invasive interventions.

Zero-heat-flux thermometry is an alternative non-invasive method that allows for continuous monitoring of core temperature. Zero-heat-flux sensors are placed on the lateral surface of the forehead and is initially warmed to equilibrate the temperature of the skin surface to the underlying core tissues. Equipped with a thermal insulator, the zero-heat-flux sensor eliminates heat loss to the environment to allow for changes in core temperature to be directly reflected by a change in skin surface temperature.[@eshraghi2014] Originally developed in the 1970s, zero-heat-flux technology was recently implemented in the the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System (3M^TM^, St Paul, MN), as a single-use, disposable sensor. This device was previously known as the SpotOn Temperature Monitoring System. Another device on the market is the T-core sensor from Dräger.

The agreement between temperature measurements from zero-heat-flux thermometers and core as well as other peripheral thermometers has been investigated in multiple studies over the last five years. The 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System, has been the zero-heat-flux device used in most of these recent studies. Appraisal and synthesis of the results would aid clinicians in deciding the appropriate circumstances in which this particular zero-heat-flux thermometer may be used. We aimed to determine if 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System has clinically acceptable accuracy and precision relative to established core and peripheral temperature measurement devices. Accuracy is defined as the average difference between temperature measurements from the zero-heat-flux and comparator device and precision as the variance (standard deviation) in the differences.

@@

A systematic review was conducted. The primary comparison for this review was zero-heat-flux temperature measured using the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System versus temperature measured from a core site, which we defined as temperature measured at either an arterial, esophageal, bladder or rectal site. Secondary comparisons were made between temperatures measured using the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System and temperatures taken at peripheral sites.

@@

Observational studies that reported temperature measurements from a 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System and comparator thermometer were included. Studies involving a case control design were excluded due to potential for overestimation of the intervention performance. Studies were excluded if conducted on non-human subjects or outside of a clinical healthcare setting. No publication date restrictions were applied. Published conference abstracts were included if there was enough information reported to appraise the quality of the study. There were no language restrictions applied during the search.

@@

Published studies were found by searching Medline and EMBASE to January 2020. The Cochrane-recommended search strategy combining terms for the ‘target condition’ and ‘index test’ was used.[@mann2012should] This search strategy is an efficient approach for systematic reviews of diagnostic test accuracy studies.[@preston2015improving] We also conducted forward citation searching, by using Google Scholar to search the citations of the first article published on the accuracy of the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System.[@eshraghi2014] The search strategies used for each database are available in the appendix. Selection of studies was undertaken independently by two reviewers.

@@

The objective for the meta-analysis was to estimate the population limits of agreement between temperature measurements from the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System and established comparator thermometers. A framework for meta-analysis of Bland-Altman method comparison studies based on limits of agreement approach was used.[@tipton2017framework]  This method was selected because it parallels the approach used in primary Bland-Altman studies, whereby an estimate is generated for the pooled limits of agreement in the population (not just in the samples studied). The ‘population limits of agreement’ is more broad than the limits of agreement commonly reported in the meta-analyses of Bland-Altman studies.[@tipton2017framework] Pooled limits of agreement were calculated using $\delta\pm2\sqrt{\sigma2+\tau2}$, where $\delta$ is the average bias across studies, $\sigma2$ is the average within-study variation in differences and $\tau2$ is the variation in bias across studies.

@@

Prior to conducting the meta-analyses, the results from each study were converted into a standard format, with bias meaning comparator thermometer temperature measurement minus the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System measurement in degrees Celsius (°C). In several studies, results were reported for multiple groups of participants, therefore in the meta-analysis each of these groups was treated as a separate 'comparison'.[@Boisson_2018; @Dahyot_Fizelier_2017; @Schell_Chaple_2018; @M_kinen_2016] Other studies reported multiple sets of results, whereby analyses were conducted between zero-heat-flux and various comparator devices used on the same participant. These instances were also treated as a separate 'comparison' if the comparator devices were a part of separate meta-analysis groups.[@West_2019; @Pesonen_2018; @M_kinen_2016; @Iden_2015] One study reported intraoperative, postoperative and overall results for the same participants[@eshraghi2014]. Only the paired measurements from the overall results were included in the main and low risk bias analyses, leaving paired measurements exclusively from the intraoperative and postoperative timepoints to be included in respective meta-analyses subgroups.

@@

Sixteen studies were included (Figure 1). Two studies reported only in abstract form were not included and assigned as 'studies awaiting classification' because there was insufficient information provided.[@cullen2015measuring; @VendrellJorda2016]

Detailed characteristics of included studies is displayed in Table 1. The primary comparison of the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System versus core temperature measurements (eg, arterial, bladder, esophageal or rectal) consisted of `r primary$Comparisons` comparisons from `r primary$Studies` individual studies.[@Boisson_2018; @Carvalho_2019; @Dahyot_Fizelier_2017; @eshraghi2014; @G_mez_Romero_2019; @idei2016; @Jack_2019; @Kato_2015; @kollmann2019validation; @M_kinen_2016; @Morettini_2019; @Pesonen_2018; @Schell_Chaple_2018; @Tachibana_2019; @G_mez_Romero_2019] In total, data from `r primary$Participants` participants with `r format(primary$Measurements, scientfic=F)` paired measurements were included in this comparison (two studies did not report the total number of measurements included in their analysis). The sensitivity analysis for the primary comparison with only studies that were assessed as low risk of bias across all domains included `r low_risk$Comparisons` comparisons from `r low_risk$Studies` studies that enrolled `r low_risk$Participants` participants with `r low_risk$Measurements` paired measurements (two studies did not report the total number of measurements included in their analysis). There were `r ICU$Studies` studies that compared zero-heat-flux to core temperature measurements in ICU patients, comprising `r ICU$Comparisons` comparisons with `r ICU$Measurements` measurements from `r ICU$Participants` participants. Another `r OT$Studies` studies were included that compared zero-heat-flux to core temperature measurements in patients undergoing surgery, comprising `r OT$Comparisons` comparisons with `r OT$Measurements` measurements from `r OT$Participants` participants. Nasopharyngeal thermometers were used as the comparator devices in `r NPA$Studies` studies, with `r NPA$Measurements` paired measurements from `r NPA$Participants` participants. Sublingual thermometers were used as the comparator devices in `r SL$Studies` studies that reported `r SL$Measurements` paired measurements from `r SL$Participants` participants. Supplementary information presented in the study [webpage](https://zhf-info.netlify.app) provides comprehensive summaries of the studies included in each of these meta-analyses.

All comparisons reported adherence to the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System manufacturer instructions and placed the sensor on the forehead of the participants. One included study also reported results for comparisons where the zero-heat-flux thermometer was not placed on the forehead.[@Tachibana_2019] We did not include these comparisons in our meta-analysis.

Risk of bias assessments for each study are presented in Figure 2. High risk of bias associated with patient selection for 13 (43%) comparisons from 9 (56%) studies, conduct of zero-heat-flux and comparator measurements in 10 (33%) comparisons from 8 (50%) studies and 12 (40%) comparisons 9 (30%) studies, respectively (mostly due to zero-heat-flux measurements being taken with knowledge of the comparator measurements and vice versa) and participant flow for 13 (43%) studies. In 19 (50%) comparisons from 8 (50%) studies, the authors had declared a conflict of interest or receipt of funding or equipment from the manufacturer of the zero-heat-flux device under evaluation.

### Primary comparison: 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System versus core thermometers

Table 1 presents results of the primary, sensitivity and all subgroup analyses for the primary comparison as well as secondary comparisons. The pooled estimate for the mean bias between the 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System and core temperature measurements was `r primary$bias_mean`°C. Population limits of agreement, which take into consideration the between-study heterogeneity and sampling error, were wide, spanning from `r primary$CI_L_rve`°C to `r primary$CI_U_rve`°C (`r primary$Measurements` measurements; `r primary$Participants` participants; `r primary$Studies` studies). The amount of between-study heterogeneity is displayed graphically in the density plot in Figure 3. The quality of evidence for the primary comparison was downgraded to low quality due to concerns about study limitations and inconsistency. Population limits of agreement for the sensitivity analysis restricted to studies rated as having low risk of bias across all the domains of the QUADAS-2 were similar to the primary analysis (`r low_risk$Measurements` measurements; `r low_risk$Participants` participants; `r low_risk$Studies` studies). The mean bias was again `r low_risk$bias_mean`°C with population limits of agreement spanning from `r low_risk$CI_L_rve`°C to `r low_risk$CI_U_rve`°C. A further sensitivity analysis excluding studies that received industry funding revealed population limits of agreement that were considerably wider than the primary analysis (`r conflict$CI_L_rve`°C to `r conflict$CI_U_rve`°C). The evidence rating for this sensitivity analysis was downgraded to low quality though, due to concerns about study limitations, inconsistency and imprecision.

We conducted two subgroup analyses for the primary comparison according to the clinical setting in which the study was conducted. In the subset of studies conducted in the ICU, the mean bias was `r ICU$bias_mean`°C with population limits of agreement between `r ICU$CI_L_rve`°C and `r ICU$CI_U_rve` °C. The GRADE rating for quality of evidence was downgraded to very low quality due to study limitations, inconsistency and imprecision. In the subset of studies that evaluated the use of 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System during surgery found a mean bias of `r OT$bias_mean`°C with population limits of agreement from `r OT$CI_L_rve`°C to `r OT$CI_U_rve`°C. The quality of evidence was rated as moderate quality due to concerns about study limitations.

### Secondary comparisons: 3M^TM^ Bair Hugger^TM^ Temperature Monitoring System versus peripheral thermometers
